---
title: "Community Flooding Sensitivity Analaysis"
author: "James Margrove"
date: "2 Dezember 2016"
output: html_document
---

Remove everything from the working environment 
```{r}
rm(list = ls())
```

Loading required packages 
```{r, message=FALSE, warning = FALSE}
require(Hmisc)
require(INLA)
require(car)
require(foreach)
require(raster)
```

Upload data, remove unknown species, drop levels and check the data.frame 
```{r}
data<- read.table("Spelev.txt",header = T)
data <- subset(data, sp != "Unkn" & mort != 1 & DBH >50) # remove unknown species
data <- droplevels(data) 
str(data)
```

Cutting the elevation gradient into bands with equal numbers of individuals in. Number of band we want to use = 12; 
```{r}
expected_n_bands <- 12
data$felv25 <- cut2(data$elev, g = expected_n_bands, m = 100)
bands <- cut2(data$elev, g = expected_n_bands, m = 100, onlycuts = TRUE)
bands[c(1,length(bands))] <- c(0,88.4) # set the min and max elevation. 
length(bands)-1 # check number of bands.
```

Take only the 16 speceis with a flooding sensivity calculated from the seedling water inundation experiment 
```{r}
SP <- c("Dry","Pmal","Ptom","Sacu","Sbec","Sfal","Sgib","Slep","Smac","Smec","Spar","Spau","Ssem","Ssmi","Swal","Sxan")
data <- data[data$sp %in% SP,]
data <- droplevels(data)
```

###To calculate the typical flooding sensitivity of an individual within each band we used INLA-SPDE. 

Preparing the mesh. Prior analysis have suggested that this mesh size is a good compormise between precision and computation time.  
```{r}
### Coordinates
coords <- cbind(data$X50N,data$Y50N)
### Mesh
mesh1 <- inla.mesh.2d(coords, max.edge = c(100, 200), 
                   cutoff =100, 
                   offset=c(200,400))
### Matern Field
spde <- inla.spde2.matern(mesh1)
### A matrix
A <- inla.spde.make.A(mesh1, coords)
```

Combining the data into a stack object. 
```{r}
stk1 <- inla.stack(tag = "stk1",
                   data = list(fsen = data$fsen), # response
                   A = list(A,1),
                   effects = list(list(i = 1:spde$n.spde),
                                  data.frame(int = 1,
                                             felv25 = data$felv25)))
```

Running the INLA-SPDE model 
```{r, warning = FALSE, message = FALSE}
model1 <- inla(fsen ~ -1 + felv25 + f(i, model = spde), 
               family = "gaussian", 
               data = inla.stack.data(stk1), 
               control.predictor = list(A=inla.stack.A(stk1), compute = TRUE),
               control.fixed = list(expand.factor.strategy = "inla"))
```

###To make a map of the flooding sensitivity bands 
Import the map and make sure all the parameters are correct
```{r}
r <- raster("Sepilok160_Qgis.tif") #This is the sepilok, map
e <- extent(603945.651,605788.486,  647133.509,  648054.815) # extent
plot160c <- crop(x = r, e) # placing the map correctly 
max(plot160c@data@values) 
plot160c <- plot160c-min(plot160c@data@values)
plot160c <- (plot160c/max(plot160c@data@values))*(88.4)
rVals <- plot160c@data@values # the values of the raster 
senPlot <- plot160c # rename the plot senPlot 
```

Defining each elevation band from the raster and change the elements in that band to teh appropreate typical flooding sensitivity. 
```{r}
pfsen <- model1$summary.fixed[,1] # The flooding sensitivity values for each band 
max(pfsen)/min(pfsen)*100 # the % increase in flooding sensitivity
```

Using a loop to put the flooding calculated sensitivity values at the correct location
```{r}
for(i in 1:length(bands)-1){
cVals <- which(rVals >= bands[i] & rVals <= bands[i+1])
senPlot@data@values[cVals] <- pfsen[i]}
```

Using ggplot to plot the raster
```{r}
ggSen <- as.data.frame(senPlot, xy = TRUE) # converting the raster to a data.frame 
colnames(ggSen) <- c("x","y","z")
### Blue lagoon colours 
ggplot(ggSen) + geom_raster(aes(x, y, fill=z)) + 
  scale_fill_gradient(low="#191654", high="#43C6AC", name = "", 
                      breaks=c(0.22, 0.25, 0.28, 0.31, 0.34)) + 
  coord_fixed()
```

```{r}
#writeRaster(senPlot, "senPlot.tif", format = "GTiff")
```



